Motional Quantum Ground-State Cooing of a Single Sodium Atom

### 2017-05-29 presentations/na_cool_paper/paper.tex
[presentations/na_cool_paper] Figure 3ab, 4ab
[presentations/na_cool_paper] Figure 3cd, 4cd

\begin{figure*}
  \includegraphics[height=5.2cm]{imgs/Na_RSC_schematic.pdf}
  \includegraphics[height=4.5cm]{sequence.pdf}
  \caption{(A) Energy levels and schematics of Raman sideband cooling.
    The Raman transitions has a one photon detuning $\Delta=25GHz$ from the D2 line.
    We use D1 light with $\sigma^-$ polarization to repump atom out of $|F=2,m_F=-1\rangle$
    state to minimize heating on the atom in $|F=2,m_F=-2\rangle$ state.
    (B) Geometry and polarizations of the Raman and optical pumping beams relative to the
    optical tweezer and bias magnetic field.
    (C) Schematics of the cooling sequence. The tweezer is switching at $3MHz$ to
    reduce light shift during optical pumping. Each cooling cycle consists of $8$ pulses.
    The four axial pulses are addressing two neighboring cooling orders.
    The two pulses in each radial directions are either addressing two neighbering cooling orders
    or having different length on the first order when most of the population are below $n=3$
    towards then end of the cooling sequence.
    \label{f-setup}}
\end{figure*}
\begin{figure}[b]
  \includegraphics[width=8.5cm]{imgs/fig2_raman_op.pdf}
  \caption{Matrix elements and heating probabilities as a function of motional state.
    The range plotted covers $99\%$ of the initial thermal distribution.
    (A) Matrix elements for Raman transition in the axial direction showing diviation from
    $\sqrt{n}$ scaling and multiple minimums for different sideband orders.
    (B) Heating probability during the optical pumping step for all three axis.
    Due to the large Lamb-Dicke parameter,
    there is a high probability of heating especially in the axial direction.
    \label{f-ld}}
\end{figure}
\begin{figure*}
  \includegraphics[width=8cm]{imgs/spectrum_r.png}
  \includegraphics[width=8cm]{imgs/temp/fit_20170409_r2_0_ba.png}
  \caption{(A) (B) \label{f-radial}}
\end{figure*}
\begin{figure*}
  \includegraphics[width=8cm]{imgs/spectrum_a1.png}
  \includegraphics[width=8cm]{imgs/temp/fit_20170409_a1_0_ba.png}
  \caption{(A) (B) \label{f-axial}}
\end{figure*}

@@

As mentioned previously, in order to perform Raman sideband cooling efficiently and
minimize the heating during the optical pumping, we need to have a low initial temperature and
a small Lamb-Dicke parameter. However, since the Lamb-Dicke parameter $\eta$ is inversely
proportional to $\sqrt{m}$ (???) where $m$ is the mass of the atom, it is larger for sodium
given the same trap depth (???).
With $45mW$ of power at the focus of the trap, we measured a trapping frequency of
$\{\omega_1,\omega_2,\omega_3\}/2\pi = \{67, 420, 580\}\ \text{kHz}$
which corresponds to optical pumping Lamb-Dicke parameters of
$\{\eta^{OP}_1,\eta^{OP}_2,\eta^{OP}_3\} = \{???, ???, ???\}$.
Moreover, due to the unresolved(?) hyperfine structure in the $3^2P_{3/2}$ manifold,
the sub-Doppler cooling in sodium is also less efficient and we start the
Raman sideband cooling with a initial temperature of $40\mu K$. Combined with the high Lamb-Dicke
parameters, this gives us a initial effective optical pumping Lamb-Dicke parameters of
$\{\eta^{OP}_{1eff},\eta^{OP}_{2eff},\eta^{OP}_{3eff}\} = \{???, ???, ???\}$
As a result, there is a very high probability of heating during the optical pumping step
(figure \ref{f-ld}B) causing a $30\%$ average heating probability during optical pumping
in the weakly confined axial direction. Fortunately, the high Lamb-Dicke parameters also
provide us tools to overcome this issue. From our geometry, the Lamb-Dicke parameters for
Raman transitions are $\{\eta^R_{1},\eta^R_{2},\eta^R_{3}\} = \{???, ???, ???\}$. As shown in
figure \ref{f-ld}A, the high Raman Lamb-Dicke parameters causes a strong coupling to higher orders
of cooling sidebands, especially for high motional states.
This enables us to cool atoms in high motional states by driving on high order Raman sidebands,
removing more motional energy in a single cooling pulse and offsetting the effect of
strong heating. Since the heating probability is higher for high motional states,
cooling on these high order sidebands can greatly suppress the high heating during
optical pumping during the initial cooling. Since the coupling strength of different orders
do not reach their minimums at the same time, using multiple orders of motional sidebands
for cooling also avoids accumulation of population near the coupling minimum of a particular
order, which improves the overall efficiency of the cooling process.

@@

Finally, the deep trap that is needed to trap and image single sodium atom also creates
a very large AC Stark shift in the excited state. For the trap depth we are using, the
light shift in the center of the trap is as large as $300MHz$.
In additional to creating a large and position dependent detuning for the optical pumping light,
it also mixes the excited state state hyperfine levels,
affecting the branching ratio and increasing the number of photons needed during optical pumping.
Similar to our loading and imaging process [???], we solve this issue by switching the
trapping light at $3MHz$ during the whole cooling sequence.
Due to the large light shift, the optical pumping is effectively off when the trap light is on.
Since the atom can only be addressed by the optical pumping light when the trap light is off,
the effect of light shift on optical pumping fidelity is suppressed.

@@

The axial direction is much less confined and therefore has a much higher initial effective
Lamb-Dicke parameter of $\eta^{OP}_{eff1}\approx3.3$.
The line in figure \ref{f-axial}A shows the Raman spectrum in the axial direction
before applying Raman sideband cooling where we can see clearly resolved Raman cooling sidebands
up to the eighth order suggesting that we have many motional states populated in this direction.
Therefore, we starts our cooling sequence by driving Raman transitions on the eighth order axial
cooling sideband. After applying the cooling sequence identical to the one we use to cool
the radial directions, the spectrum is shown as the blue line in figure \ref{f-axial}A.
All of the high order ($\geqslant2$) axial cooling sidebands disappeared and the first order
cooling sideband is also strongly suppressed.
The ground state probability calculated from this spectrum is $????$.
We can also use Rabi flopping on the carrier and heating sideband to varify this result
similar to the radial direction. In this case, we see very good agreement on the carrier
(figure \ref{f-axial}B) but there is additional decoherence on the axial first order
heating sideband (figure \ref{f-axial}C).
We believe this decoherence is caused by technical noise in our setup which
produces the strongest effect on this transition due to its slow Rabi frequency.
The decoherence time scale agree with the magnetic field fluctuation we measured which produces
a Zeeman shift of around $5kHz$.

>>>>

\begin{figure*}
  \includegraphics[height=5.2cm]{imgs/Na_RSC_schematic.pdf}
  \includegraphics[height=4.5cm]{sequence.pdf}
  \caption{(A) Energy levels and schematics of Raman sideband cooling.
    The Raman transitions has a one photon detuning $\Delta=25GHz$ from the D2 line.
    We use D1 light with $\sigma^-$ polarization to repump atom out of $|F=2,m_F=-1\rangle$
    state to minimize heating on the atom in $|F=2,m_F=-2\rangle$ state.
    (B) Geometry and polarizations of the Raman and optical pumping beams relative to the
    optical tweezer and bias magnetic field.
    (C) Schematics of the cooling sequence. The tweezer is switching at $3\text{MHz}$ to
    reduce light shift during optical pumping. Each cooling cycle consists of $8$ pulses.
    The four axial pulses are addressing two neighboring cooling orders.
    The two pulses in each radial directions are either addressing two neighbering cooling orders
    or having different length on the first order when most of the population are below $n=3$
    towards then end of the cooling sequence.
    \label{f-setup}}
\end{figure*}
\begin{figure}[b]
  \includegraphics[width=8.5cm]{imgs/fig2_raman_op.pdf}
  \caption{Matrix elements and heating probabilities as a function of motional state.
    The range plotted covers $99\%$ of the initial thermal distribution.
    (A) Matrix elements for Raman transition in the axial direction showing diviation from
    $\sqrt{n}$ scaling and multiple minimums for different sideband orders.
    (B) Heating probability during the optical pumping step for all three axis.
    Due to the large Lamb-Dicke parameter,
    there is a high probability of heating especially in the axial direction.
    \label{f-ld}}
\end{figure}
\begin{figure*}
  \includegraphics[height=4cm]{imgs/spectrum_r.pdf}
  \includegraphics[height=4cm]{imgs/rabi_flop_r3_0.png}
  \includegraphics[height=4cm]{imgs/rabi_flop_r3_p1.png}
  \caption{(A) (B) \label{f-radial}}
\end{figure*}
\begin{figure*}
  \includegraphics[height=4cm]{imgs/spectrum_a1.pdf}
  \includegraphics[height=4cm]{imgs/rabi_flop_a1_0.png}
  \includegraphics[height=4cm]{imgs/rabi_flop_a1_p1.png}
  \caption{(A) (B) \label{f-axial}}
\end{figure*}

@@

As mentioned previously, in order to perform Raman sideband cooling efficiently and
minimize the heating during the optical pumping, we need to have a low initial temperature and
a small Lamb-Dicke parameter. However, since the Lamb-Dicke parameter $\eta$ is inversely
proportional to $\sqrt{m}$ (???) where $m$ is the mass of the atom, it is larger for sodium
given the same trap depth (???).
With $45mW$ of power at the focus of the trap, we measured a trapping frequency of
$\{\omega_1,\omega_2,\omega_3\}/2\pi = \{67, 420, 580\}\ \text{\text{kHz}}$
which corresponds to optical pumping Lamb-Dicke parameters of
$\{\eta^{OP}_1,\eta^{OP}_2,\eta^{OP}_3\} = \{???, ???, ???\}$.
Moreover, due to the unresolved(?) hyperfine structure in the $3^2P_{3/2}$ manifold,
the sub-Doppler cooling in sodium is also less efficient and we start the
Raman sideband cooling with a initial temperature of $40\mu K$. Combined with the high Lamb-Dicke
parameters, this gives us a initial effective optical pumping Lamb-Dicke parameters of
$\{\eta^{OP}_{1eff},\eta^{OP}_{2eff},\eta^{OP}_{3eff}\} = \{???, ???, ???\}$
As a result, there is a very high probability of heating during the optical pumping step
(figure \ref{f-ld}B) causing a $30\%$ average heating probability during optical pumping
in the weakly confined axial direction. Fortunately, the high Lamb-Dicke parameters also
provide us tools to overcome this issue. From our geometry, the Lamb-Dicke parameters for
Raman transitions are $\{\eta^R_{1},\eta^R_{2},\eta^R_{3}\} = \{???, ???, ???\}$. As shown in
figure \ref{f-ld}A, the high Raman Lamb-Dicke parameters causes a strong coupling to higher orders
of cooling sidebands, especially for high motional states.
This enables us to cool atoms in high motional states by driving on high order Raman sidebands,
removing more motional energy in a single cooling pulse and offsetting the effect of
strong heating. Since the heating probability is higher for high motional states,
cooling on these high order sidebands can greatly suppress the high heating during
optical pumping during the initial cooling. Since the coupling strength of different orders
do not reach their minimums at the same time, using multiple orders of motional sidebands
for cooling also avoids accumulation of population near the coupling minimum of a particular
order, which improves the overall efficiency of the cooling process.

@@

Finally, the deep trap that is needed to trap and image single sodium atom also creates
a very large AC Stark shift in the excited state. For the trap depth we are using, the
light shift in the center of the trap is as large as $300\text{MHz}$.
In additional to creating a large and position dependent detuning for the optical pumping light,
it also mixes the excited state state hyperfine levels,
affecting the branching ratio and increasing the number of photons needed during optical pumping.
Similar to our loading and imaging process [???], we solve this issue by switching the
trapping light at $3\text{MHz}$ during the whole cooling sequence.
Due to the large light shift, the optical pumping is effectively off when the trap light is on.
Since the atom can only be addressed by the optical pumping light when the trap light is off,
the effect of light shift on optical pumping fidelity is suppressed.

@@

The axial direction is much less confined and therefore has a much higher initial effective
Lamb-Dicke parameter of $\eta^{OP}_{eff1}\approx3.3$.
The line in figure \ref{f-axial}A shows the Raman spectrum in the axial direction
before applying Raman sideband cooling where we can see clearly resolved Raman cooling sidebands
up to the eighth order suggesting that we have many motional states populated in this direction.
Therefore, we starts our cooling sequence by driving Raman transitions on the eighth order axial
cooling sideband. After applying the cooling sequence identical to the one we use to cool
the radial directions, the spectrum is shown in figure \ref{f-axial}B.
All of the high order ($\geqslant2$) axial cooling sidebands disappeared and the first order
cooling sideband is also strongly suppressed.
The ground state probability calculated from this spectrum is $????$.
We can also use Rabi flopping on the carrier and heating sideband to varify this result
similar to the radial direction. In this case, we see very good agreement on the carrier
(figure \ref{f-axial}B) but there is additional decoherence on the axial first order
heating sideband (figure \ref{f-axial}C).
We believe this decoherence is caused by technical noise in our setup which
produces the strongest effect on this transition due to its slow Rabi frequency.
The decoherence time scale agree with the magnetic field fluctuation we measured which produces
a Zeeman shift of around $5\text{kHz}$.
